Manuscript

### 2018-11-08 manuscript.Rmd

@@

## Make G2P lineage plotly
```{r}
make_G2Plineage_plotly <- function(lineage_name) {
  use_data <- filter(g2p_druggenelin_all_summ, group_general_lineage_name == lineage_name)
  use_data$Dataset <- factor(use_data$Dataset, levels = c("CCLE", "CTRP", "GDSC"))
  ticker <- 0

@@

@@

## Drugscreen G2P evidence

### By gene

@@

```{r, eval=FALSE, results = 'hide'}
lapply(as.character(unique(g2p_druggenelin_all_summ$group_general_lineage_name)), make_G2Plineage_plotly)

@@

### Gene expression

```{r}
drugscreen_ptmuts <- rbind(ccle_data_ptmuts, ctrp_data_ptmuts, gdsc_data_ptmuts)
```

```{r}
summ <- ccle_data_ptmuts %>% count(group_general_lineage_name, Drug_Gene)
```

```{r}
PearsonByLineage <- function(lineage, dataset) {
  use_data <- filter(dataset, group_general_lineage_name == lineage)
  if(nrow(use_data) == 0) {
    use_data <- NA
  }
  # use_data_summ <- use_data %>% count(Drug_Gene)
  # use_data_summ <- filter(use_data_summ, n >= 3)
  # use_data <- filter(use_data, Drug_Gene %in% as.character(unique(use_data_summ$Drug_Gene)))
  # 
  # pearson_res <- use_data %>% group_by(Drug_Gene) %>% summarize(Pearson_Corr = cor.test(y = AUC, x = RPKM, method = "pearson", use = "complete.obs")$estimate, Pearson_Pval = cor.test(y = AUC, x = RPKM, method = "pearson", use = "complete.obs")$p.value)
  # pearson_res <- merge(g2p_druggene_p_tally, pearson_res, all.y = TRUE)
  # pearson_res$group_general_lineage_name <- lineage

  return(use_data)
}

# test <- PearsonByLineage("breast cancer", ccle_data_ptmuts)

# ccle_signif_pearson <- lapply(as.character(unique(ccle_data_ptmuts$group_general_lineage_name)), PearsonByLineage, dataset = ccle_data_ptmuts)
```

```{r}
PearsonPlotlyByLineage <- function(lineage) {

  ticker <- 0

  use_data_ccle <- filter(use_data, Dataset == "CCLE")
  use_data_ccle_summ <- use_data_ccle %>% count(InG2P)
  if(nrow(use_data_ccle_summ) <= 1) {
    ccle_ttest_res <-  NA
    ticker <- ticker + 1
    }
  else if(use_data_ccle_summ[use_data_ccle_summ$InG2P == "Yes",]$nn < 2) { ccle_ttest_res <-  NA }
  else if(use_data_ccle_summ[use_data_ccle_summ$InG2P == "Yes",]$nn >= 2) {
    ccle_lin_p_ttest <- t.test(use_data_ccle$p ~ use_data_ccle$InG2P)
    ccle_ttest_res <- paste0("t = ", round(ccle_lin_p_ttest$statistic, 3), ", df = ", round(ccle_lin_p_ttest$parameter, 3), ", p = ", round(ccle_lin_p_ttest$p.value, 3))
  }

  use_data_ctrp <- filter(use_data, Dataset == "CTRP")
  use_data_ctrp_summ <- use_data_ctrp %>% count(InG2P)
  if(nrow(use_data_ctrp_summ) <= 1) {
    ctrp_ttest_res <-  NA
    ticker <- ticker + 1
    }
  else if(use_data_ctrp_summ[use_data_ctrp_summ$InG2P == "Yes",]$nn < 2) { ctrp_ttest_res <-  NA }
  else if(use_data_ctrp_summ[use_data_ctrp_summ$InG2P == "Yes",]$nn >= 2) {
    ctrp_lin_p_ttest <- t.test(use_data_ctrp$p ~ use_data_ctrp$InG2P)
    ctrp_ttest_res <- paste0("t = ", round(ctrp_lin_p_ttest$statistic, 3), ", df = ", round(ctrp_lin_p_ttest$parameter, 3), ", p = ", round(ctrp_lin_p_ttest$p.value, 3))
  }

  use_data_gdsc <- filter(use_data, Dataset == "GDSC")
  use_data_gdsc_summ <- use_data_gdsc %>% count(InG2P)
  if(nrow(use_data_gdsc_summ) <= 1) {
    gdsc_ttest_res <-  NA
    ticker <- ticker + 1
    }
  else if(use_data_gdsc_summ[use_data_gdsc_summ$InG2P == "Yes",]$nn < 2) { gdsc_ttest_res <-  NA }
  else if(use_data_gdsc_summ[use_data_gdsc_summ$InG2P == "Yes",]$nn >= 2) {
    gdsc_lin_p_ttest <- t.test(use_data_gdsc$p ~ use_data_gdsc$InG2P)
    gdsc_ttest_res <- paste0("t = ", round(gdsc_lin_p_ttest$statistic, 3), ", df = ", round(gdsc_lin_p_ttest$parameter, 3), ", p = ", round(gdsc_lin_p_ttest$p.value, 3))
  }

  g2p_druggenelin_all_summ_text <- data.frame(Dataset = c("CCLE", "CTRP", "GDSC"), Test_Results_p = c(ccle_ttest_res, ctrp_ttest_res, gdsc_ttest_res))
  # use_data_summ <- use_data %>% count(InG2P)
  # if(nrow(use_data_ccle_summ) <= 1) { ttest_res <- NA }
  # else if(use_data_summ[use_data_summ$InG2P == "Yes",]$nn < 2) { ttest_res <- NA }
  # else if(use_data_summ[use_data_summ$InG2P == "Yes",]$nn >= 2) {
  #   ttest <- t.test(use_data$p ~ use_data$InG2P)
  #   ttest_res <- paste0("t = ", round(ttest$statistic, 3), ", df = ", round(ttest$parameter, 3), ", p = ", round(ttest$p.value, 3))
  # }
}
```

@@

plot_pearson_box <- ggplot(data = g2p_druggene_pearson) +
  facet_wrap(~ Dataset) +
  geom_violin(mapping = aes(color = InG2P, y = Pearson_Corr, x = InG2P), alpha = 0.5) +
  geom_boxplot(mapping = aes(color = InG2P, y = Pearson_Corr, x = InG2P), alpha = 0.75, width = 0.25) +
  coord_flip() +
  labs(x = "Level A G2P\nassociation") +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

@@

>>>>

## Pearson correlation by lineage
```{r}
PearsonByLineage <- function(lin, dataset) {
  use_data <- filter(dataset, group_general_lineage_name == lin)
  use_data <- filter(use_data, !is.na(RPKM))
  use_data_summ <- use_data %>% count(Drug_Gene)
  use_data_summ <- filter(use_data_summ, n >= 3)
  use_data_filt <- filter(use_data, Drug_Gene %in% as.character(unique(use_data_summ$Drug_Gene)))
  if(nrow(use_data_filt) > 0) {
    pearson_res <- use_data_filt %>% group_by(Drug_Gene) %>% summarize(Pearson_Corr = cor.test(y = AUC, x = RPKM, method = "pearson", use = "complete.obs")$estimate, Pearson_Pval = cor.test(y = AUC, x = RPKM, method = "pearson", use = "complete.obs")$p.value)
    pearson_res <- merge(g2p_druggene_p_tally, pearson_res, all.y = TRUE)
    pearson_res$group_general_lineage_name <- lin
    return(pearson_res)
  }
}
```

@@

## Make G2P lineage p-value plotly
```{r}
MutPlotlyByLineage <- function(lineage_name) {
  use_data <- filter(g2p_druggenelin_all_summ, group_general_lineage_name == lineage_name)
  use_data$Dataset <- factor(use_data$Dataset, levels = c("CCLE", "CTRP", "GDSC"))
  ticker <- 0

@@

## Make G2P lineage gene expression plotly
```{r}
GexpPlotlyByLineage <- function(lineage) {
  use_data <- filter(signif_pearson_all_lin, group_general_lineage_name == lineage)
  use_data$Dataset <- factor(use_data$Dataset, levels = c("CCLE", "CTRP", "GDSC"))
  use_data$Pearson_Pvalue <- paste0(round(use_data$Pearson_Pval, 3), "\nDrug_Gene: ", use_data$Drug_Gene)
  ticker <- 0

  use_data_ccle <- filter(use_data, Dataset == "CCLE")
  use_data_ccle_summ <- use_data_ccle %>% count(InG2P)
  if(nrow(use_data_ccle_summ) <= 1) {
    ccle_ttest_res <-  NA
    ticker <- ticker + 1
    }
  else if(use_data_ccle_summ[use_data_ccle_summ$InG2P == "Yes",]$nn < 2) { ccle_ttest_res <-  NA }
  else if(use_data_ccle_summ[use_data_ccle_summ$InG2P == "Yes",]$nn >= 2) {
    ccle_lin_ttest <- t.test(use_data_ccle$Pearson_Corr ~ use_data_ccle$InG2P)
    ccle_ttest_res <- paste0("t = ", round(ccle_lin_ttest$statistic, 3), ", df = ", round(ccle_lin_ttest$parameter, 3), ", p = ", round(ccle_lin_ttest$p.value, 3))
  }

  use_data_ctrp <- filter(use_data, Dataset == "CTRP")
  use_data_ctrp_summ <- use_data_ctrp %>% count(InG2P)
  if(nrow(use_data_ctrp_summ) <= 1) {
    ctrp_ttest_res <-  NA
    ticker <- ticker + 1
    }
  else if(use_data_ctrp_summ[use_data_ctrp_summ$InG2P == "Yes",]$nn < 2) { ctrp_ttest_res <-  NA }
  else if(use_data_ctrp_summ[use_data_ctrp_summ$InG2P == "Yes",]$nn >= 2) {
    ctrp_lin_ttest <- t.test(use_data_ctrp$Pearson_Corr ~ use_data_ctrp$InG2P)
    ctrp_ttest_res <- paste0("t = ", round(ctrp_lin_ttest$statistic, 3), ", df = ", round(ctrp_lin_ttest$parameter, 3), ", p = ", round(ctrp_lin_ttest$p.value, 3))
  }

  use_data_gdsc <- filter(use_data, Dataset == "GDSC")
  use_data_gdsc_summ <- use_data_gdsc %>% count(InG2P)
  if(nrow(use_data_gdsc_summ) <= 1) {
    gdsc_ttest_res <-  NA
    ticker <- ticker + 1
    }
  else if(use_data_gdsc_summ[use_data_gdsc_summ$InG2P == "Yes",]$nn < 2) { gdsc_ttest_res <-  NA }
  else if(use_data_gdsc_summ[use_data_gdsc_summ$InG2P == "Yes",]$nn >= 2) {
    gdsc_lin_ttest <- t.test(use_data_gdsc$Pearson_Corr ~ use_data_gdsc$InG2P)
    gdsc_ttest_res <- paste0("t = ", round(gdsc_lin_ttest$statistic, 3), ", df = ", round(gdsc_lin_ttest$parameter, 3), ", p = ", round(gdsc_lin_ttest$p.value, 3))
  }

  y_value <- ifelse(ticker == 3, 10, max(use_data$n) + 1)
  x_value <- (max(use_data$Pearson_Corr) + min(use_data$Pearson_Corr)) / 2

  summ_text <- data.frame(Dataset = c("CCLE", "CTRP", "GDSC"), Test_Results = c(ccle_ttest_res, ctrp_ttest_res, gdsc_ttest_res))

  plot_point <- ggplot(data = use_data) +
    facet_wrap(~ Dataset, drop = FALSE) +
    scale_y_continuous(breaks = pretty_breaks()) +
    geom_point(mapping = aes(label = Pearson_Pvalue, color = InG2P, x = Pearson_Corr, y = n), alpha = 0.5, position = position_jitter(width = 0, height = 0.15)) +
    geom_text(data = summ_text, mapping = aes(x = x_value, y = y_value, label = Test_Results), nudge_y = 1) +
    labs(x = "Gene-drug association p-value between mutation groups", y = "Number of level A G2P associations") +
    theme(legend.position = "none", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank())

  plot_box <- ggplot(data = use_data) +
    facet_wrap(~ Dataset, drop = FALSE) +
    geom_violin(mapping = aes(color = InG2P, y = Pearson_Corr, x = InG2P), alpha = 0.5) +
    geom_boxplot(mapping = aes(color = InG2P, y = Pearson_Corr, x = InG2P), alpha = 0.5, width = 0.5) +
    coord_flip() +
    labs(x = "Level A G2P\nassociation") +
    theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

  plot <- subplot(plot_box, plot_point, nrows = 2, titleX = TRUE, titleY = TRUE) %>% layout(title = lineage, margin = list(t = 75, l = 75), xaxis5 = list(title = "Pearson correlation coefficient between gene-drug AUC and gene expression"), yaxis = list(title = "Level A G2P\nassociation", domain = c(0.8, 1)), yaxis2 = list(title = "Number of level A G2P associations", domain = c(0, 0.79)))

  saveWidgetFix(widget = plot, file = paste0("./plots_18Q3/manuscript/plotlys_pearson_by_lineage/", lineage, "_g2p_evidence_pearson_plotly.html"))
}
```

@@

## Drugscreen: Mutation status G2P evidence

### Lineage-agnostic

@@

```{r, eval=FALSE, results = 'hide'}
lapply(as.character(unique(g2p_druggenelin_all_summ$group_general_lineage_name)), MutPlotlyByLineage)

@@

## Drugscreen: Gene expression G2P evidence

### Lineage-agnostic

@@

plot_pearson_box <- ggplot(data = g2p_druggene_pearson) +
  facet_wrap(~ Dataset) +
  geom_violin(mapping = aes(color = InG2P, y = Pearson_Corr, x = InG2P), alpha = 0.5) +
  geom_boxplot(mapping = aes(color = InG2P, y = Pearson_Corr, x = InG2P), alpha = 0.5, width = 0.25) +
  coord_flip() +
  labs(x = "Level A G2P\nassociation") +
  theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

@@

```{r, eval=FALSE}
ccle_lineages4pearson <- as.character(unique(ccle_data_ptmuts$group_general_lineage_name))[!is.na(as.character(unique(ccle_data_ptmuts$group_general_lineage_name)))]
ccle_signif_pearson <- lapply(ccle_lineages4pearson, PearsonByLineage, dataset = ccle_data_ptmuts)
ccle_signif_pearson_all_lin <- rbindlist(ccle_signif_pearson, use.names = TRUE)
ccle_signif_pearson_all_lin$Dataset <- "CCLE"

ctrp_lineages4pearson <- as.character(unique(ctrp_data_ptmuts$group_general_lineage_name))[!is.na(as.character(unique(ctrp_data_ptmuts$group_general_lineage_name)))]
ctrp_signif_pearson <- lapply(ctrp_lineages4pearson, PearsonByLineage, dataset = ctrp_data_ptmuts)
ctrp_signif_pearson_all_lin <- rbindlist(ctrp_signif_pearson, use.names = TRUE)
ctrp_signif_pearson_all_lin$Dataset <- "CTRP"

gdsc_lineages4pearson <- as.character(unique(gdsc_data_ptmuts$group_general_lineage_name))[!is.na(as.character(unique(gdsc_data_ptmuts$group_general_lineage_name)))]
gdsc_signif_pearson <- lapply(gdsc_lineages4pearson, PearsonByLineage, dataset = gdsc_data_ptmuts)
gdsc_signif_pearson_all_lin <- rbindlist(gdsc_signif_pearson, use.names = TRUE)
gdsc_signif_pearson_all_lin$Dataset <- "GDSC"

signif_pearson_all_lin <- rbind(ccle_signif_pearson_all_lin, ctrp_signif_pearson_all_lin, gdsc_signif_pearson_all_lin)
signif_pearson_all_lin$n <- ifelse(is.na(signif_pearson_all_lin$n), 0, signif_pearson_all_lin$n)
signif_pearson_all_lin$InG2P <- ifelse(signif_pearson_all_lin$n != 0, "Yes", "No")
```

```{r, results='hide', eval=FALSE}
lapply(as.character(unique(signif_pearson_all_lin$group_general_lineage_name)), GexpPlotlyByLineage)
```
